Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-09 article/article.tex
Commit 2024-06-09 18:17

For the detection of CE cases we established a baseline of irradiance
above which we characterized each data point as an enhancement event and
calculated the over irradiance (OIR). The OIR is defined as the
irradiance difference of the measured one-minute \(\text{GHI}_i\) from
the \(\text{GHI}_\text{ref}\) corresponding to cloud-free atmosphere.
First, we tested two simple approaches for the determination of
\(\text{GHI}_\text{ref}\): The Haurwitz's model \citep{Haurwitz1945},
which is a simple clear sky model and was already adjusted and applied
to our data \citep{Natsis2023}, and the total solar irradiance (TSI) at
the top of the atmosphere, adjusted for the Sun-Earth distance. We have
tested both cases by using an appropriate relative threshold and/or an
additional constant offset. The initial results, showed that we can
detect a big portion of the CE events. These results were helpful, as
they are independent from unknown factors, and helped us to establish
some criteria to further improve the CE identification. It was evident,
by inspecting the daily plot of irradiance, that changes on the
atmospheric conditions introduced numerous false positive and false
negative results. To produce a more accurate reference, we had to take
into account more factors that effect the clear sky radiation. So we
used a radiation transfer model in order to include the effects of ADO
and the water vapors.

\hypertarget{modeled-clear-sky-irradiance}{%
\subsection{Modeled clear Sky
Irradiance}\label{modeled-clear-sky-irradiance}}

@@

We approximated the expected clear sky GHI by using the Libradtran
radiation transfer model \citep{Emde2016}, a similar approach, was also
used by \citet{Vamvakas2020} for creating a clear sky reference for
studding cloud enhancement events. Main factors responsible for the
attenuation of the broadband downward solar radiation in the atmosphere
are the aerosols and the water vapors. Because of the lack of
observational data for the whole period, we used some long term
climatological data to recreate the whole period. Fortunately, our site
participates in the Aerosol Robotic Network (AERONET)
\citep{Giles2019, Buis1998}, as there is in operation a Cimel photometer
since 2003, collocated with the CM-21 pyranometer. The mean monthly AOD
on different wavelengths are provided by AERONET, along with the
equivalent water column height in the atmosphere.

For the atmospheric characteristics, we iterated on combinations of AOD
at \(500\,\text{nm}\) (\(\tau_{500\text{nm}}\)) with additional offsets
of \(\pm1\) and \(\pm2\sigma\), and water column (\(w\)), also with
offsets of \(\pm1\) and \(\pm2\sigma\). We applied them on two
atmospheric profiles, from the Air Force Geophysics Laboratory (AFGL).
The ``AFGL atmospheric constituent profile, midlatitude summer''
(afglms) and the ``AFGL atmospheric constituent profile, midlatitude
winter'' (afglmw) \citep{Anderson1986}. --\textgreater{}

>>>>

For the detection of CE cases we established a baseline of irradiance
above which we characterized each data point as an enhancement event and
calculated the over irradiance (OIR). The OIR is defined as the
irradiance difference of the measured one-minute \(\text{GHI}_i\) from
the \(\text{GHI}_\text{ref}\) corresponding to cloud-free atmosphere.
First, we tested two simple approaches for the determination of
\(\text{GHI}_\text{ref}\): The Haurwitz's model \citep{Haurwitz1945},
which is a simple clear sky model and was already adjusted and applied
to our data \citep{Natsis2023}, and the total solar irradiance (TSI) at
the top of the atmosphere, adjusted for the Sun-Earth distance. We have
tested both methods by using an appropriate relative threshold and/or an
additional constant offset. The initial results showed that we can
detect a big portion of the CE events. However, by inspecting the daily
plots of irradiance it became evident that changes in atmospheric
conditions introduced numerous false positive and false negative
results. The main reason for these discrepancies is the variability of
the effects of aerosols and water vapor which were not taken into
account in the two simple methods. To produce a more representative
reference we included the effects of these factors through variables
using a radiative transfer model (RTM). The applied methodology is
discussed in section\nobreakspace{}\ref{rtmcs}

\hypertarget{rtmcs}{%
\subsection{Modeled clear Sky Irradiance}\label{rtmcs}}

@@

We approximated the expected clear sky \(\text{GHI}_\text{ref}\) with
the radiative transfer model uvspec, part of libRadtran
\citep{Emde2016}, similarly to the approach used by
\citet{Vamvakas2020}. In uvspec we used the solar spectrum of
\citet{Kurucz1994} in the range \(280\) to \(2500\,\text{nm}\), the
radiative trasfer solver ``disort'' in ``pseudospherical'' geometry and
the ``LOWTRAN'' gas parameterization. The model was run for a range of
variables in order to create a look up table (LUT) for the estimation of
the cloud-free reference irradiance for each individual observation of
our dataset. In this context, the model was run for SZAs in the range
\(10\) -- \(90^\circ\) with a step of \(0.2^\circ\) and for the
atmospheric profiles of the Air Force Geophysics Laboratory
\citep{Anderson1986} midlatitude summer and midlatitude winter,
representative of the warm and cold seasons.

Main factors responsible for the attenuation of the broadband downward
solar radiation under cloud free atmospheres are aerosols and water
vapor. At Thessaloniki, such measurements are available since 2003 from
the Cimel sun photometer that belongs to the Aerosol Robotic Network
(AERONET) \citep{Giles2019, Buis1998}. The mean monthly AOD on different
wavelengths are provided by AERONET, along with the equivalent water
column height in the atmosphere.

For the atmospheric characteristics, we iterated on combinations of AOD
at \(500\,\text{nm}\) (\(\tau_{500\text{nm}}\)) with additional offsets
of \(\pm1\) and \(\pm2\sigma\), and water column (\(w\)), also with
offsets of \(\pm1\) and \(\pm2\sigma\).
